---
title: "Identifying persons of interest in Enron dataset"
output: html_document
---

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(gridExtra)
```


```{r load_data, cache=TRUE}
enron <- read.csv("final_project_dataset.csv", na.strings = "NaN")
```

## Financials

```{r, message=FALSE, fig.width=10}
ggplot(enron, aes(salary, fill=poi)) + 
  geom_histogram(alpha=0.5, position="identity") +
  scale_fill_brewer(palette="Set1")
```

Some kind of outlier in the top graph.

```{r}
enron %>%
  arrange(desc(salary)) %>%
  select(name, salary) %>%
  head()
```

Let's remove the 'TOTAL' data.

```{r}
enron <- enron %>%
  filter(name != "TOTAL")
```

Replot the salary data.

```{r, message=FALSE, fig.width=10}
ggplot(enron, aes(salary, fill=poi)) + 
  geom_histogram(alpha=0.5, position="identity") +
  scale_fill_brewer(palette="Set1")
```

Looks better.

Let's plot all the financial features and see how they compare between POIs and
non-POIs.

```{r, messages=FALSE, warning=FALSE, fig.width=10, fig.height=40}
financial_features=c(
  "salary",
  "deferral_payments",
  "total_payments",
  "exercised_stock_options",
  "bonus",
  "restricted_stock",
  "restricted_stock_deferred",
  "total_stock_value",
  "expenses",
  "other",
  "director_fees",
  "deferred_income",
  "long_term_incentive"
)

plots = lapply(financial_features, function(feature) {
  return(ggplot(enron, aes_string(feature, fill="poi")) + 
           geom_histogram(alpha=0.5, position="identity"))
})

plots$ncol=1

do.call(grid.arrange, plots)
```

deferral_payments - noticeably different, worth a closer look, weird to see a 
value less than zero.

total_payments - outlier worth a closer look

exercised_stock_options - seems to be a feature worth trying

bonus - worth a try, not clear how useful it will be

restricted_stock - not interesting

restricted_stock_deferred - not interesting

total_stock_value - maybe
expenses - not interesting

other - outlier worth a closer look

director_fees - not interesting

deferred_income - maybe

long_term_incentive - not interesting

```{r, messages=FALSE, fig.width=10, fig.height=40}
enron %>%
  arrange(deferral_payments) %>%
  select(name, deferral_payments) %>%
  head()

enron %>%
  filter(name=="BELFER ROBERT")
```

The data for this person seems to be shifted columnwise. 
Let's just ignore this row.

```{r}
enron <- enron %>%
  filter(name!="BELFER ROBERT")
```

```{r, messages=FALSE, fig.width=10, fig.height=40}
enron %>%
  arrange(desc(total_payments)) %>%
  select(name, total_payments) %>%
  head()
```

Doesn't look like total_payments would do well to tell apart POIs.

```{r, messages=FALSE, fig.width=10, fig.height=40}
enron %>%
  arrange(desc(other)) %>%
  select(name, other) %>%
  head()
```

Outlier is Kenneth Lay again. If we ignore him, this feature doesn't seem to a 
good indicator.

Let's try using deferral_payments, exercised_stock_options, bonus, 
total_stock_value, deferred_income from financial features.

What if we look at fraction of payments/stock by type?

```{r, fig.width=10, fig.height=20}
enron <- enron %>%
  mutate(salary_pct=salary/total_payments,
         bonus_pct=bonus/total_payments,
         long_term_incentive_pct=long_term_incentive/total_payments,
         deferral_payments_pct=deferral_payments/total_payments,
         other_pct=other/total_payments,
         expenses_pct=expenses/total_payments,
         director_fees_pct=director_fees/total_payments)
financial_features=c(
  "salary_pct",
  "bonus_pct",
  "long_term_incentive_pct",
  "deferral_payments_pct",
  "other_pct",
  "expenses_pct",
  "director_fees_pct"
)

plots = lapply(financial_features, function(feature) {
  return(ggplot(enron, aes_string(feature, fill="poi")) + 
           geom_histogram(alpha=0.5, position="identity"))
})
plots$ncol=1

do.call(grid.arrange, plots)
```

Overall, financial features appear to be weak indicators of POIs.

## E-mails

What about e-mail features?

to_messages
shared_receipt_with_poi
from_messages
from_this_person_to_poi
from_poi_to_this_person

```{r, messages=FALSE, fig.width=10, fig.height=20}
g1 <- ggplot(enron, aes(to_messages, fill=poi)) + geom_histogram(alpha=0.5, position="identity")
g2 <- ggplot(enron, aes(shared_receipt_with_poi, fill=poi)) + geom_histogram(alpha=0.5, position="identity")
g3 <- ggplot(enron, aes(from_messages, fill=poi)) + geom_histogram(alpha=0.5, position="identity")
g4 <- ggplot(enron, aes(from_this_person_to_poi, fill=poi)) + geom_histogram(alpha=0.5, position="identity")
g5 <- ggplot(enron, aes(from_poi_to_this_person, fill=poi)) + geom_histogram(alpha=0.5, position="identity")

grid.arrange(g1, g2, g3, g4, g5, ncol=1)
```

None of these features look particularly interesting. 
What if we look at the percentage of emails from/to POI?

```{r}
enron <- enron %>%
  mutate(from_this_person_to_poi_pct = from_this_person_to_poi / from_messages,
         from_poi_to_this_person_pct = from_poi_to_this_person / to_messages)
```

```{r}
ggplot(enron, aes(from_this_person_to_poi_pct, fill=poi)) + geom_histogram(alpha=0.5, position="identity")
ggplot(enron, aes(from_poi_to_this_person_pct, fill=poi)) + geom_histogram(alpha=0.5, position="identity")
```

from_this_person_to_poi_pct might be useful, but again none of the email 
features appear to be great indicators of POIs.

